---
title: "ERA5: Download via R"
author: "Simeon Lisovski"
date: "2021-05-26"
tags:
- R
- era5
- wind
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You can downlaod wind (horizontal) u und (vertical) v vectors on a global scale for different pressure levels (altitudes) from the ERA-5 dataset.
Batch download requests can be done with python code that can be run via RStudio.

First, you need to create an account with [ECWMF](https://www.ecmwf.int/). 

Next, you need to have some kind of Python compiler installed on you pc/mac. See https://rstudio.github.io/reticulate/ for explanations. This link may also help to get the minimum required python stuff installed (https://docs.conda.io/en/latest/miniconda.html).

You need to following packages:

```{r packages, message=FALSE, warning=FALSE}
library(reticulate)
library(glue)
```

Set your Python environment.

```{r, warning=FALSE, message=FALSE, eval = FALSE}
use_virtualenv("~/Documents/Virtualenv", required = TRUE) ## here, I have installed python 3
```


## Test

As a test, downlad the global wind dataset for a specific day and two specific pressure levels:

```{r, eval = FALSE}
date    <- as.POSIXct("2020-10-08")
plevels <- glue_collapse(glue("'{c(650, 1000)}'"), ", ")

py_run_string(glue(
          "
          import cdsapi
          c = cdsapi.Client()
          
          c.retrieve(
          'reanalysis-era5-pressure-levels',
          {{
              'product_type': 'reanalysis',
              'format': 'netcdf',
              'variable': [
                  'u_component_of_wind', 'v_component_of_wind',
              ],
              'pressure_level': [{plevels},],
              'year': '{format(date, '%Y')}',
              'month': '{format(date, '%m')}',
              'day': '{format(date, '%d')}',
              'time': ['12:00',],
          }},
          'ERA5_wind_{format(date, '%Y')}_{format(date, '%m')}_{format(date, '%d')}.nc') 
          "
          ))
```

## Load into R

```{r, echo=FALSE}
date    <- as.POSIXct("2020-10-08")
plevels <- glue_collapse(glue("'{c(650, 1000)}'"), ", ")
```


```{r, message=FALSE, warning=FALSE}
library(raster)
library(ncdf4)
library(sf)
library(rnaturalearthdata)

## read data for first pressure level [[1]]
uRaster <- brick(glue("ERA5_wind_{format(date, '%Y')}_{format(date, '%m')}_{format(date, '%d')}.nc"), varname=  "u")[[1]] 
vRaster <- brick(glue("ERA5_wind_{format(date, '%Y')}_{format(date, '%m')}_{format(date, '%d')}.nc"), varname=  "v")[[1]]

speedRaster <- rotate(sqrt((uRaster^2)+(vRaster^2))) ## rotate to have -180 to 180 degrees (otherwhise 0 to 360)
# plot(speedRaster)
# plot(rnaturalearthdata::coastline50, add = T)
```
